---
title: "Untitled"
author: "Ashfaq Ali"
date: "10/15/2020"
output: html_document
---


## Gene/Protein annotations from biomaRt.

```{r}
library(biomaRt)
ensembl = useMart("ensembl",dataset="mmusculus_gene_ensembl")
filters = listFilters(ensembl)
filters[1:10,]
attributes = listAttributes(ensembl)
attributes[1:15,]
searchAttributes(mart = ensembl, pattern = "uniprot")

genes <- geneID # this would correspond to gene ID from your data
proteins <- proteinID  # this would correspond to protein ID from your data
genes

hgnc_symbol

genes2 <-getBM(attributes=c('hgnc_symbol', 'entrezgene_id', 'description'), 
      filters = 'external_gene_name ',  ## uniprotswissprot
      values = genes, 
      mart = ensembl
        )
select(ensembl)
```

### Load gene sets

For mor details about the gene sets, please consult (https://www.gsea-msigdb.org/gsea/msigdb). 

```{r}
browseVignettes("clusterProfiler")
```



```{r, message=FALSE, warning=FALSE}
library(msigdbr)
library(clusterProfiler)

m_df <- msigdbr(species = "Mus musculus") ## download the gene sets from msigdb

hm_gsig <-m_df %>% filter(gs_cat=="H") %>% dplyr::select(gs_name, gene_symbol) # Hallmark genesets

C5_gsig <-m_df %>% filter(gs_cat=="C5") %>% dplyr::select(gs_name, gene_symbol) # Go terms

C2_gsig <-m_df %>% filter(gs_cat=="C2" & gs_subcat == "CP:KEGG") %>% dplyr::select(gs_name, gene_symbol) # Curated gene sets (Kegg pathways)
C3_gsig <-m_df %>% filter(gs_cat=="C3" & gs_subcat == "TFT:GTRD") %>% dplyr::select(gs_name, gene_symbol) # Motifs (Trancription factor targets)
C7_gsig <-m_df %>% filter(gs_cat=="C7") %>% dplyr::select(gs_name, gene_symbol) # C7 (immunologic signatures, 4872 gene sets)

```

### Gene Set Enrichment analyses

```{r}
Adhesion_geneset <- list()
W1 <- AgeMain$logFC # Onne needs to provide that rank of the genes 

names(W1)<- rownames(AgeMain)

W1 <- sort((abs(W1)), decreasing = TRUE) 
W1 <- W1[!duplicated(names(W1))]
#mutate(rank = rank(log2FoldChange,  ties.method = "random")) 
Ageing_Patways$Hallmark <-GSEA(W1, TERM2GENE = hm_gsig, exponent = 1, 
                            #nPerm = 10000, 
             pvalueCutoff = 0.2, minGSSize = 3, maxGSSize = 600,  verbose = FALSE, by="fgsea", eps = 0) %>% as.data.frame() %>% tibble() %>% mutate(enrichment= ifelse(NES > 0, "Up-regulated", "Down-regulated"), Description = stringr::str_replace_all(ID, c("HALLMARK_"= "", "_"=" ")))

Ageing_Patways$kegg <-GSEA(W1, TERM2GENE = C2_gsig, exponent = 1, 
                        #nPerm = 10000, 
             pvalueCutoff = 0.2, minGSSize = 3, maxGSSize = 600,  verbose = FALSE, by="fgsea", eps = 0) %>% as.data.frame() %>% tibble() %>% mutate(enrichment= ifelse(NES > 0, "Up-regulated", "Down-regulated"), Description = stringr::str_replace_all(ID, c("KEGG_"= "", "_"=" ")))

Ageing_Patways$Regulation <-GSEA(W1, TERM2GENE = C3_gsig, exponent = 1, 
                              #nPerm = 10000, 
                              pvalueCutoff = 0.2, minGSSize = 3, maxGSSize = 600,  verbose = FALSE, by="fgsea", eps = 0) %>% as.data.frame() %>% tibble() %>% mutate(enrichment= ifelse(NES > 0, "Up-regulated", "Down-regulated"))

Ageing_Patways$GO <-GSEA(W1, TERM2GENE = C5_gsig, exponent = 1, 
                      #nPerm = 10000,  
                      pvalueCutoff = 0.2, minGSSize = 3, maxGSSize = 600,  verbose = FALSE, by="fgsea", eps = 0) %>% as.data.frame() %>% tibble() %>% mutate(enrichment= ifelse(NES > 0, "Up-regulated", "Down-regulated"), Description = stringr::str_replace_all(ID, c("GO_"= "", "_"=" ")))


Ageing_Patways$Immunological <-GSEA(W1, TERM2GENE = C7_gsig, exponent = 1, 
                                 #nPerm = 10000, 
             pvalueCutoff = 0.2, minGSSize = 3, maxGSSize = 600,  verbose = FALSE, by="fgsea", eps = 0) %>% as.data.frame() %>% tibble() %>% mutate(enrichment= ifelse(NES > 0, "Up-regulated", "Down-regulated"))


# ßwritexl::write_xlsx(Ageing_Patways, "../Results/Gene_sets.xlsx")

  
```



_______________________________________
##GO 

```{r}
gene <- names(geneList)[abs(geneList) > 2]
gene.df <- bitr(gene, fromType = "ENTREZID",
        toType = c("ENSEMBL", "SYMBOL"),
        OrgDb = org.Hs.eg.db)
head(gene.df)
```


```{r}
 ggo <- groupGO(gene     = gene,
               OrgDb    = org.Hs.eg.db,
               ont      = "CC",
               level    = 3,
               readable = TRUE)

head(ggo)
```


```{r}



ego <- enrichGO(gene          = gene,
                universe      = names(geneList),
                OrgDb         = org.Hs.eg.db,
                ont           = "CC",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)

```

```{r}
ego2 <- enrichGO(gene         = gene.df$ENSEMBL,
                OrgDb         = org.Hs.eg.db,
                keyType       = 'ENSEMBL',
                ont           = "CC",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.05)
```


```{r}
ego2 <- setReadable(ego2, OrgDb = org.Hs.eg.db)
```

geneList contains three features:

numeric vector: fold change or other type of numerical variable
named vector: every number has a name, the corresponding gene ID
sorted vector: number should be sorted in decreasing order
If you import your data from a csv file, the file should contains two columns, one for gene ID (no duplicated ID allowed) and another one for fold change.

```{r}




ego3 <- gseGO(geneList     = geneList,
              OrgDb        = org.Hs.eg.db,
              ont          = "CC",
              nPerm        = 1000,
              minGSSize    = 100,
              maxGSSize    = 500,
              pvalueCutoff = 0.05,
              verbose      = FALSE)
```

